---
title: 'Reproducible Research: Week 2 Course Project'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Code for reading in the dataset

```{r}
data <- read.csv("activity.csv", header = TRUE, sep = ',', na.strings="NA", stringsAsFactors=T)
```


## Changing the formating of date column
```{r}
library(lubridate)
data$date <- ymd(data$date)
```


## Histogram of the total number of steps taken each day

```{r}
library(ggplot2)
aggregate_data <- aggregate(steps ~ date, data, sum)

ggplot(aggregate_data, aes(x = steps)) +
        geom_histogram(binwidth = 1000) +
        labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")

```

## Mean and median number of steps taken each day

```{r}

mean_steps <- mean(aggregate_data$steps, na.rm = TRUE)

print(mean_steps)

median_steps <- median(aggregate_data$steps, na.rm = TRUE)

print(median_steps)

```


## Time series plot of the average number of steps taken


```{r}

aggregate_data_interval <- aggregate(steps ~ interval, data, mean)


ggplot(aggregate_data_interval, aes(x=interval, y=steps)) +
        geom_line(color = "firebrick")

```


## Time series plot of the average number of steps taken


```{r}

aggregate_data_interval[which.max(aggregate_data_interval$steps),]


```



## Code to describe and show a strategy for imputing missing data


```{r}

data_full <- data
nas <- is.na(data_full$steps)
avg_interval <- tapply(data_full$steps, data_full$interval, mean, na.rm=TRUE, simplify=TRUE)
data_full$steps[nas] <- avg_interval[as.character(data_full$interval[nas])]

```


## Histogram of the total number of steps taken each day after missing values are imputed

```{r}
aggregate_data_full <- aggregate(steps ~ date, data_full, sum)


ggplot(aggregate_data_full, aes(x = steps)) +
        geom_histogram(binwidth = 1000) +
        labs(title = "Histogram of the total number of steps taken each day after missing values are imputed
", x = "Steps per day", y = "Frequency")

```

## Mean and Median after imputing the values
```{r}

mean_impute <- mean(aggregate_data_full$steps)
print(mean_impute)

median_impute <- median(aggregate_data_full$steps)
print(median_impute)

```


## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
library(dplyr)
data_full <- mutate(data_full, weektype = ifelse(weekdays(data_full$date) == "Saturday" | weekdays(data_full$date) == "Sunday", "weekend", "weekday"))
data_full$weektype <- as.factor(data_full$weektype)


aggregate_data_full_weekday <- aggregate(steps ~ interval + weektype, data_full, mean)


panel_plot <- ggplot(aggregate_data_full_weekday, aes(x=interval, y=steps, color = weektype)) +
        geom_line() +
        facet_wrap(~weektype, ncol = 1, nrow=2)
print(panel_plot)

```

